---
title: "Plotting"
author: "Emi Carlevaro"
date: today
date-meta: 15/12/2025
editor: 
  mode: source
format: 
  html:
    code-fold: true
    code-tools: true
    self-contained: true
    page-layout: full
    toc: true
params:
  OUTFOLDER: "/"
  NCORES: 8
bibliography: [book.bib, packages.bib]
biblio-style: apalike
---

```{r}
library(dplyr)
library(magrittr)
library(iStabi)
library(plotly)

```

```{r}
Results <- readRDS('Results.rds')
Explored <- Results$Explored
```


## PLOTTING

THe S-test follows a Chi-Square distribution with **`r KZMKX`** degrees of freedom
(# of instruments =`r NCOL(Z)` - # of included instruments =`r NCOL(X)`

### Critical values
Critical values for $qLL-\tilde{S}$ for `r KZ` instruments ($k = `r KZ`$), from table VII Suplementary material, p3 Magnusson Mavroeidis (2014).


THe S-test follows a Chi-Square distribution with **`r NCOL(Z)-NCOL(X)`** degrees of freedom (# of instruments = ` `r NCOL(Z)` ` - \# of included instruments = ` `r NCOL(X)` `

`Explored` has `r NROW(Explored)` rows.

Critical values for $qLL-\tilde{S}$ for **`r KZ`** instruments ($k = `r KZ`$), from table VII Suplementary material, p3 Magnusson Mavroeidis (2014).

The number of *fixed* identified parameters is **`r NFIXPARAMS`**.

The number of *estimated* parameters under the null is **`r NSIPARAMS`**.


```{r}
# (C)ritical (V)alues//test//# of instruments
# CVs$qLLStab$2
CVs <- list('qLLStab' = list(
                tibble('sig' = c(0.1, 0.05, 0.01, 0),
                      'value' = c(0, 7.17, 8.36, 11.10)),
                tibble('sig' = c(0.1, 0.05, 0.01, 0),
                      'value' = c(0, 12.79, 14.30, 17.58)),
                # the value is the floor: an statistic between 0 and 17.58 will have a sig of 0.1, and between 17.58 and +\infty will have 0
                tibble('sig' = c(0.1, 0.05, 0.01, 0),
                    'value' = c(0, 18.14, 19.95, 23.51)) ))
# Select appropiate critical values¯
CV <- CVs$qLLStab[[KZ]]
```

### Data for plotting

```{r}
Sdf =  2
pExplored <- filter(Explored, !if_all(.fns=is.na)) %>%
  # findInterval: the second vector gives the FLOOR value for each interval
  mutate(., 
         'qLLStab_sigAt' = CV$sig[ findInterval(qLLStab, CV$value,left.open=TRUE) ],
         'S_pval' = round(1-pchisq(S, df=Sdf), digits=2),
         'pointSize' = 0, 
         'inQllSet' = if_else(qLLStab_sigAt>=0.10, 1, 0))
# This should be TRUE (the qll value is for 3 instruments)
NROW(filter(pExplored, qLLStab<=23.51)) == NROW(filter(pExplored, qLLStab_sigAt>=0.01))

saveRDS(pExplored, 'pExplored.rds')

```

### Grid for plotting
Force the starting values to be a multiple of the step size. This avoids problems and is clear what the true parameter space is.

### Plots
True parameter vector $ \color{blue}{\theta} = `r thetaStar` $

#### Global search
Scatterplot for all searched values (useful for the global search)

##### S plot
```{r}
# INTERACTIVE PLOTs ----
#### S-plot ----
pSbase <- plot_ly(data = filter(pExplored, S<4.60517), 
                 x=~gammaF, 
                 y=~lambda,
                 color=) %>%
  layout(title = 'S-set for the New Keyneasian Phillips Curve (90% confidence)',
         xaxis = list(title=plotly::TeX('\\gamma_F'), 
                        tickvals=seq(from=0, to=1, by=0.10),
                        range = c(0, 1),
                        fixedrange = TRUE),
           yaxis = list(title=plotly::TeX('\\lambda'),  
                        tickvals=seq(from=0.001, to=0.2, by=0.05),
                        range = c(0.001, 0.2),
                        fixedrange = TRUE)) %>%
  add_trace(x = 0.591,y = 0.015,
    type = "scatter",   mode = "markers",
    marker = list(symbol = "square",  size = 12,  color = "red" ),
    inherit = FALSE,
    showlegend = FALSE) %>%
  config(mathjax = "cdn")

pSbase %>%
  add_trace(type='scatter', mode='markers', marker = list(color='blue'))

```

##### qLL-Stab plot
```{r}
# INTERACTIVE PLOTs ----
#### S-plot ----
pSbase <- plot_ly(data = filter(pExplored, qLLStab <= 18.14), 
                 x=~gammaF, 
                 y=~lambda,
                 color=) %>%
  layout(title = 'S-set for the New Keyneasian Phillips Curve (90% confidence)',
         xaxis = list(title=plotly::TeX('\\gamma_F'), 
                        tickvals=seq(from=0, to=1, by=0.10),
                        range = c(0, 1),
                        fixedrange = TRUE),
           yaxis = list(title=plotly::TeX('\\lambda'),  
                        tickvals=seq(from=0.001, to=0.2, by=0.05),
                        range = c(0.001, 0.2),
                        fixedrange = TRUE)) %>%
  config(mathjax = "cdn")

pSbase %>%
  add_trace(type='scatter', mode='markers')

```

```{r}
#### qLLStab plot ----
pQllExplored <- filter(pExplored, qLLStab_sigAt >= 0.1) %>%
  # Finding edges of the plot
  group_by(., gammaF) %>% 
  mutate('edge_min' = if_else(lambda == min(lambda), TRUE, FALSE),
         'edge_max' = if_else(lambda == max(lambda), TRUE, FALSE)) %>%
  ungroup()

pQllbase <- plot_ly(data = pQllExplored, 
                  x=~gammaF, y=~lambda, color=~qLLStab) %>%
  layout(title = 'qLL-S set for New Keyneasian Phillips Curve (90% confidence)',
         scene=list(
           xaxis = list(title=plotly::TeX('\\gamma_f'), 
                        tickvals=seq(from=PARAMS_CFG$minVal[1], to=PARAMS_CFG$maxVal[1],
                                     by=PARAMS_CFG$stepSize[1])),
           yaxis = list(title=plotly::TeX('\\lambda'),  
                        tickvals=seq(from=PARAMS_CFG$minVal[2],           to=PARAMS_CFG$maxVal[2], 
                                     by=PARAMS_CFG$stepSize[2])))) %>%
  config(mathjax = "cdn")

pQllbase %>%
  add_trace(type='scatter', mode='markers')

```

```{r}

pBase <- ggplot(data=pQllExplored, mapping=aes(x=gammaF, y=lambda)) +
  scale_x_continuous(name=TeX(r"(Respose to expected inflation ($\gamma_f$))"), limits=c(-0.1, 1.1), 
                     breaks=seq(from=-0.1, to=1.1, by=PARAMS_CFG$stepSize[1]), expand = expand_scale(mult=0.01)) +
  scale_y_continuous(name=TeX(r'(Respose to interest rate ($\lambda$))'), limits=c(-0.3, 0.3), 
                     breaks=seq(from=-0.3, to=0.3, by=4*(PARAMS_CFG$stepSize[2])), expand = expand_scale(mult=0.01)) + 
  theme_bw() + theme(panel.grid=element_line(colour='#999999', linetype='14'), 
                     panel.grid.minor=element_line(colour='white'), 
                     text=element_text(size=10), 
                     axis.title.x = element_text(colour='red'),
                     axis.title.y = element_text(colour='blue', size=rel(0.9), hjust=0.9),
                     axis.text.x = element_text(size=rel(1.25), angle=90, vjust=0.5), 
                     axis.text.y = element_text(size=rel(1.25))) +
  geom_vline(xintercept=0, colour = 'black', linetype = 'solid') +
  geom_hline(yintercept=0, colour = 'black', linetype = 'solid') +   
  # Highlightinh the TRUE value
  geom_point(aes(x=thetaStar['gammaF'], y=thetaStar['lambda']), size=3, colour='black', fill='yellow')

# Scatterplot
pBase + geom_point(aes(colour=qLLStab))
# Curvy set
pBase +
  coord_cartesian(xlim=c(-0.1, 1.1), ylim=c(-0.3, 0.3), expand=TRUE) + 
  geom_raster(fill='grey', colour='black',
            interpolate=TRUE, show.legend=TRUE, alpha=0.8) + 
  # Lower edge
  geom_path(data=filter(pQllExplored, edge_min == TRUE))  +
  # Upper edge
  geom_path(data=filter(pQllExplored, edge_max == TRUE))
# RA


```

